XRONOS: an open data infrastructure for archaeological chronometry

### 2024-01-30 paper.qmd
Render self-contained HTML

  This paper outlines the conceptual and technical infrastructure developed to realise these goals in XRONOS’ initial phases of development (2019 and 2021–2023), including a generalised data model for site and radiocarbon information, extendable to other chronometric data; an R- and Ruby-based pipeline for continuous ingestion of data from a variety of sources; continuous, semi-automated data cleaning protocols; a Ruby-on-Rails application providing a web-based frontend to the data and a REST API for programmatic access; and an R package for interfacing with the API. We believe the XRONOS framework provides more open, more reliable, and more comprehensive access to chronometric data than previously available, as well as a foundation for its continuous expansion and refinement.
copyright: 
  holder: the authors
  year: 2023
license: "CC BY"
# citation: 
#  container-title: Archaeometry
#  volume: 1
#  issue: 1
#  doi: 10.5555/12345678
funding: "This work was funded by the Swiss National Science Foundation ([SNSF Project #198152](https://data.snf.ch/grants/grant/198153))"
bibliography: references.bib
format:
  html:
    page-layout: article
    toc: true
---

>>>>

  This paper outlines the conceptual and technical infrastructure developed to realise these goals in XRONOS’ initial phases of development (2019 and 2021–2023), including a generalised data model for site and radiocarbon information, extendable to other chronometric data; an R- and Ruby-based pipeline for continuous ingestion of data from a variety of sources; continuous, semi-automated data cleaning protocols; a Ruby-on-Rails application providing a web-based frontend to the data and a REST API for programmatic access; and an R package for interfacing with the API. We believe the XRONOS framework provides more open, more reliable, and more comprehensive access to chronometric data than previously available, as well as a foundation for its continuous expansion and refinement.
copyright: 
  holder: the authors
  year: 2023
license: "CC BY"
# citation: 
#  container-title: Archaeometry
#  volume: 1
#  issue: 1
#  doi: 10.5555/12345678
funding: "This work was funded by the Swiss National Science Foundation ([SNSF Project #198152](https://data.snf.ch/grants/grant/198153))"
bibliography: references.bib
format:
  html:
    page-layout: article
    toc: true
    embed-resources: true
---
